---
title: "Final Project IS607"
author: "Marco Siqueira Campos, KaMan Chan, Sharon Morris, Talha Muhammad"
date: "12/18/2016"
output: html_document
---

# Introduction
This is the final project for IS607 Data Acquisition and Management class in the M.S. of Data Analytics program at The City University of New York. 

This is an observational study. 

$H_0 :p =0.50$ (There is no relationship between property values and crime)
$H_a :p >0.50$ (There is a relationship between property values and crime)

## Motivation
The motivation of this project is to understand the relationship between crime and property values in New York City. The majority of the project team are New York City residents and this is of interest.

The results of this study could be useful to those interested in purchasing real estate in New York City.

## The Data

The data were obtained from two sources. Crime data were collected from New York City Police Department complaint data. The crime dataset includes all valid felony, misdemeanor, and violation crimes reported to the New York City Police Department (NYPD) during 2015. These data represent criminal offenses according to New York State Penal Law definitions, not FBI Uniform Crime Report definitions, and are therefore not comparable to UCR reported crime.

Each row represents a complaint reported to NYPD. Only valid complaints are included in this release. Complaints deemed unfounded due to reporter error or misinformation are excluded from the data set, as they are not reflected in official figures nor are they considered to have actually occurred in a criminal context. Similarly, complaints that were voided due to internal error are also excluded from the data set.

Property sales data from August to November 2016 were scraped from the real estate listing site Trulia. 


#### Obtain the data

```{r}
library(dplyr)
library(plyr)
library(corrplot)
library(ggmap)
library(RMySQL)
```


```{r}
#temp <- tempfile()
#download.file("https://raw.githubusercontent.com/cunyauthor/FinalProject/master/NYPD_Complaint_Data_Historic.zip",temp)
#NYPD <- read.csv(unz(temp, "NYPD_Complaint_Data_Historic.csv"), encoding="UTF-8", na.strings=c("","NA"), stringsAsFactors = F)
#unlink(temp)

#str(NYPD)
#head(NYPD)
```


#### Scrub and explore the data

A subset of the crime data was created with only variables required for the analysis. All missing data were removed from the dataset, the resulting dataset contained 100,000+ cases. 

The resulting dataset was randomized to create to create a sample of 11,700 cases. The sample was broken into 5 smaller samples in order to add Google API to match longitude and latitude in the data to street addresses. Google API allowed 2,500 free downloads per 24 hour period.

An index was created to categorize crime and applied. The crime dataset and data from Trulia were joined into one dataset for analysis.

```{r}
#class(NYPD) #view class
#dim(NYPD)   #view dimensions
#names(NYPD) #explore the column names
#unstand the structure
#str(NYPD)  

#new <- select(NYPD, RPT_DT, OFNS_DESC, OFNS_DESC, LAW_CAT_CD, BORO_NM, ADDR_PCT_CD,
              #Latitude, Longitude, Lat_Lon) # data with required variables

#sapply(new, function(x) sum(is.na(x))) # review NAs in the data
#noNas <- na.omit(new) # remove NAs from data
#sapply(noNas, function(x) sum(is.na(x)))
#glimpse(noNas) #look at the data

#summary(noNas) #summarize the data
```


##### Create datasets with Google API addresses, crime index
```{r}
#set.seed(123) # set the seed to standartize
#nyc<-noNas[sample(nrow(noNas),458557),] # random the sample to standardize the sequence
#nyc$id<-1:458557 # create a id to check

# Create datasets with Google API
#nyc_1<-nyc[1:2400,]# 
#res <- mapply(FUN = function(lon, lat) {
        #revgeocode(c(lon, lat), output = "more")
#},
#nyc_1$Longitude, nyc_1$Latitude
#)
#res1<-rbind_all(lapply(res, as.data.frame))
#View(res1)
#nyc_1<-cbind(nyc_1,res1) # add full address to data base
#write.csv(nyc_51 file = 'nyc_1.csv', row.names = FALSE) #write to csv
#head(nyc_1)

#nyc_2<-nyc[2401:4800,]  
#res <- mapply(FUN = function(lon, lat) {
        #revgeocode(c(lon, lat), output = "more")
#},
#nyc_2$Longitude, nyc_2$Latitude
#)
#res1<-rbind_all(lapply(res, as.data.frame))
#View(res1)
#nyc_2<-cbind(nyc_2,res1) # add full address to data base
#write.csv(nyc_2, file = 'nyc_2.csv', row.names = FALSE) #write to csv
#head(nyc_2)

#nyc_3<-nyc[4801:6900,]# 2nd Dec Sharon sample # (for today)
#res <- mapply(FUN = function(lon, lat) {
        #revgeocode(c(lon, lat), output = "more")
#},
#nyc_3$Longitude, nyc_3$Latitude
#)
#res1<-rbind_all(lapply(res, as.data.frame))
#View(res1)
#nyc_3<-cbind(nyc_3,res1) # add full address to data base
#write.csv(nyc_3, file = 'nyc_3.csv', row.names = FALSE) #write to csv
#head(nyc_3)

#nyc_4<-nyc[7201:9600,]# 
#res <- mapply(FUN = function(lon, lat) {
        #revgeocode(c(lon, lat), output = "more")
#},
#nyc_4$Longitude, nyc_4$Latitude
#)
#res1<-rbind_all(lapply(res, as.data.frame))
#View(res1)
#nyc_4<-cbind(nyc_4,res1) # add full address to data base
#write.csv(nyc_4, file = 'nyc_4.csv', row.names = FALSE) #write to csv
#head(nyc_4)

#nyc_5<-nyc[9600:12000,]# 
#res <- mapply(FUN = function(lon, lat) {
        #revgeocode(c(lon, lat), output = "more")
#},
#nyc_5$Longitude, nyc_5$Latitude
#)
#res1<-rbind_all(lapply(res, as.data.frame))
#View(res1)
#nyc_5<-cbind(nyc_5,res1) # add full address to data base
#write.csv(nyc_5, file = 'nyc_5.csv', row.names = FALSE) #write to csv
#head(nyc_5)

# read the nyc data
urlfile<-"https://raw.githubusercontent.com/cunyauthor/FinalProject/master/nyc_15.csv"
nyc_1<-read.csv(url(urlfile), encoding="UTF-8", na.strings=c("","NA"), stringsAsFactors = F)

nyc_1$postal_code<-as.character(nyc_1$postal_code)

# create new column for violent crime 
index<-c("DANGEROUS WEAPONS","FELONY ASSAULT","KIDNAPPING & RELATED OFFENSES","MURDER & NON-NEGL. MANSLAUGHTER","ROBBERY","SEX CRIMES")
nyc_1$crime_v<-index[(match(nyc_1$OFNS_DESC,index))]

nyc_1$crime_v<-as.factor(nyc_1$crime_v)

# agrregate the crime data by zip code
nyc_s1<-aggregate(OFNS_DESC ~ postal_code, nyc_1, length)
nyc_s2<-aggregate(crime_v ~ postal_code, nyc_1, length)

colnames(nyc_s1) <- c("zip","freq_crime")
colnames(nyc_s2) <- c("zip","freq_violent_crime")

# joint two tables of crime data
nyc_s12<-left_join(nyc_s1, nyc_s2, "zip")

# change NA for zero 
nyc_s12[is.na(nyc_s12)]<-0

# Crime table
head(nyc_s12)

# Read data from real state 
urlfile<-"https://raw.githubusercontent.com/cunyauthor/FinalProject/master/trulia_sqft.csv"
trulia<-read.csv(url(urlfile), encoding="UTF-8", na.strings=c("","NA"), stringsAsFactors = F)

##NEW CODE
# Read the table
urlfile<-"https://raw.githubusercontent.com/cunyauthor/FinalProject/master/zip_neig.csv"
zip_neig<-read.csv(url(urlfile), encoding="UTF-8", na.strings=c("","NA"), stringsAsFactors = F)
 
#joint crime data with borough and neighborhood
 zip_neig$zip<-as.character(zip_neig$zip)
nyc_s12<-left_join(nyc_s12,zip_neig, "zip")
head(nyc_s12)
#END NEW DATA

# Statistics summary of NYC prices
summary(trulia$Median_sales_price)
summary(trulia$Price_sqft)

# Trulia table price by zip
head(trulia)

# Change zip to as.character
trulia$zip<-as.character(trulia$zip)

# Join the two tables
nyc_full<-inner_join(trulia,nyc_s12, "zip")
head(nyc_full)

# read population by ZIP frm 2010 US Census
urlfile<-"https://raw.githubusercontent.com/cunyauthor/FinalProject/master/pop_zip_census.csv"
zip_pop<-read.csv(url(urlfile), encoding="UTF-8", na.strings=c("","NA"), stringsAsFactors = F)

colnames(zip_pop) <- c("zip","pop")
zip_pop$zip<-as.character(zip_pop$zip)

# Join the two tables, add population data
nyc_full<-inner_join(nyc_full,zip_pop, "zip")

# Include crime rate data per 1000.000 habit per year by zip
nyc_full$crime_rate<-(nyc_full$freq_crime/nyc_full$pop)*100000 * 458557/nrow(nyc_1)
nyc_full$v_crime_rate<-(nyc_full$freq_violent_crime/nyc_full$pop)*100000 * 458557/nrow(nyc_1)

```


```{r}
# Database schema
source("logincredentiasl.R")
connection <- dbConnect(MySQL(), user=MySQL_Username, password=MySQL_Password)

dbSendQuery(connection, 'CREATE SCHEMA IF NOT EXISTS SYS;')
dbSendQuery(connection, 'USE SYS;')
dbSendQuery(connection, 'DROP TABLE IF EXISTS nyc_1;')
dbSendQuery(connection, 'Drop Table If Exists nyc_full;')
dbSendQuery(connection, 'Drop Table If Exists nyc_s1;')
dbSendQuery(connection, 'Drop Table If Exists nyc_s12;')
dbSendQuery(connection, 'Drop Table If Exists nyc_s2;')
dbSendQuery(connection, 'Drop Table If Exists trulia;')
dbSendQuery(connection, 'Drop Table If Exists zip_pop;')

urlfile<-"https://raw.githubusercontent.com/cunyauthor/FinalProject/master/nyc_15.csv"
nyc_1<-read.csv(url(urlfile), encoding="latin1", na.strings=c("","NA"), stringsAsFactors = F)

dbWriteTable(connection, "tbl_nyc1", nyc_1, append = TRUE, row.names = FALSE)
dbSendQuery(connection, "ALTER TABLE tbl_nyc1
            MODIFY COLUMN id varchar(100) NOT NULL,            
            MODIFY COLUMN RPT_DT varchar(100) NOT NULL,
            MODIFY COLUMN OFNS_DESC varchar(100) NOT NULL,
            MODIFY COLUMN BORO_NM varchar(100) NOT NULL,
            MODIFY COLUMN ADDR_PCT_CD varchar(100) NULL,
            MODIFY COLUMN Latitude varchar(100) NULL,
            MODIFY COLUMN Longitude varchar(100) NULL,
            MODIFY COLUMN Lat_Lon varchar(100) NULL,
            MODIFY COLUMN address varchar(200) NULL,
            MODIFY COLUMN street_number varchar(100) NULL,
            MODIFY COLUMN route varchar(100) NULL,
            MODIFY COLUMN neighborhood varchar(100) NULL,
            MODIFY COLUMN political varchar(100) NULL,
            MODIFY COLUMN administrative_area_level_2 varchar(100) NULL,
            MODIFY COLUMN administrative_area_level_1 varchar(100) NULL,
            MODIFY COLUMN country varchar(100) NULL,
            MODIFY COLUMN postal_code varchar(100) NULL,
            MODIFY COLUMN postal_code_suffix varchar(100) NULL,
            MODIFY COLUMN locality varchar(100) NULL,
            MODIFY COLUMN premise varchar(100) NULL,
            MODIFY COLUMN crime_v varchar(100) NULL
            ;")

urlfile<-"https://raw.githubusercontent.com/cunyauthor/FinalProject/master/pop_zip_census.csv"
zip_pop<-read.csv(url(urlfile), encoding="UTF-8", na.strings=c("","NA"), stringsAsFactors = F)
colnames(zip_pop) <- c("zip","pop")# rename variables
dbWriteTable(connection, "tbl_zip_pop", zip_pop, append = TRUE, row.names = FALSE)
dbSendQuery(connection, "ALTER TABLE tbl_zip_pop
            MODIFY COLUMN zip varchar(10) NOT NULL,
            MODIFY COLUMN pop varchar(50) NOT NULL
            ;")

urlfile<-"https://raw.githubusercontent.com/cunyauthor/FinalProject/master/zip_neig.csv"
zip_neig<-read.csv(url(urlfile), encoding="UTF-8", na.strings=c("","NA"), stringsAsFactors = F)

dbWriteTable(connection, "tbl_zip_neig", zip_neig, append = TRUE, row.names = FALSE)
dbSendQuery(connection, "ALTER TABLE tbl_zip_neig
            MODIFY COLUMN Borough varchar(100) NOT NULL,
            MODIFY COLUMN Neighborhood varchar(100) NOT NULL,
            MODIFY COLUMN zip varchar(10) NOT NULL
            ;")

urlfile<-"https://raw.githubusercontent.com/cunyauthor/FinalProject/master/trulia_sqft.csv"
trulia<-read.csv(url(urlfile), encoding="UTF-8", na.strings=c("","NA"), stringsAsFactors = F)

dbWriteTable(connection, "tbl_trulia", trulia, append = TRUE, row.names = FALSE)
dbSendQuery(connection, "ALTER TABLE tbl_trulia
            MODIFY COLUMN x varchar(10) NOT NULL,
            MODIFY COLUMN zip varchar(10) NOT NULL,
            MODIFY COLUMN Median_sales_price varchar(50) NOT NULL,
            MODIFY COLUMN Price_sqft varchar(50) NOT NULL
            ;")

```


#### Model the Data

```{r}
#  Correlation matrix
#nyc_full$X <- NULL
#m<-cor(subset(nyc_full, select=-c(zip)))
#corrplot(m, type="upper", order="hclust", t1.col="black",t1.srt=45)

# Scatterplot with regression line
ggplot(nyc_full,aes(y = Price_sqft, x = freq_violent_crime)) +
    geom_point() + ggtitle("Sales Price per Square Foot vs Violent Crime by ZIP") + 
    xlab("Violent Crime") + ylab("Price per Square Foot") + 
    geom_smooth(method="lm", fill=NA) +
    theme(plot.title = element_text(color="blue", size=14, face="bold"),
          axis.title.x = element_text(color="black", size=10, face="bold"),
          axis.title.y = element_text(color="black", size=10, face="bold"))

# Regression analysis
lmt<-lm(Price_sqft ~ freq_violent_crime, nyc_full)
summary(lmt)

```

#### Findings and Conclusion

The scatter plot shows there is not a strong a relationship between price per square
and violent crimes. 

Our regression analysis shows for every violent crime the sales price per square foot drops by $32.90. The amount in price per square foot drops can vary by $11.20. There is a relationship between sales price per square foot and violent crime in New York City. Thus we reject the null hypothesis.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r tableau}
```
<iframe src="https://public.tableau.com/profile/ka.man.chan#!/"
 width="650" height="1100">
 </iframe>
```


#### References
The definition for violent crime was created using two existing standards NCVS [Bureau of Justice Statistics National Crime Victimization Survey](https://www.bjs.gov/index.cfm?ty=dcdetail&iid=245) and the UCR [Federal Bureau of Investigation's Uniform Crime Report](https://ucr.fbi.gov)

**Trulia real state**
https://www.trulia.com/home_prices/New_York/New_York-heat_map/city_by_neighborhood/ALP/nh/
 
**NYC Opendata – City crime data**
https://data.cityofnewyork.us/Public-Safety/Historical-New-York-City-Crime-Data/hqhv-9zeg
 
**US Census 2010**
http://www.census.gov/2010census/data/
 
**New York State**
https://www.health.ny.gov/statistics/cancer/registry/appendix/neighborhoods.htm
 
**Violent crime definition**
https://en.wikipedia.org/wiki/Violent_crime
   
**Google API**
https://developers.google.com/maps/pricing-and-plans/
 
**R ggmap (Google maps) package**                                https://cran.r-project.org/web/packages/ggmap/ggmap.pdf
 